Quality Definitions of Qualities

### 2019-12-04 StateOfPractice/QDefOfQualities/QDefOfQualities.tex
Reformatted tex file for consistency, 80 character width

\begin{defn}
	A measure of the ease of installation \citep{smith2015comparing}.\wss{Where
	does this definition come from? Please add a citation.  Is it from
	\citet{McCallEtAl1977}? I thought so originally, but after looking into it, I
	don't think so anymore.}
	\oo{@smith you added the definition in here but I guess it's from here}
\end{defn}

@@

\wss{The rationale is missing.  Please add.} \oo{Done}
The definition captures the cost or effort - human or machine, and considers not just the initial installation but uninstallation and re-installation which are all related to installation activities.
\oo{Done installability}
\subsection{\notdone{Correctness} \oo{owner}}

@@

\begin{defn}
    Testability is the degree to which a requirement is stated in terms that permits establishment of test criteria and performance of tests to determine whether those criteria have been met \citep{IEEEStdGlossarySET1990}
\end{defn}
\begin{defn}
	Code possesses the characteristic testability to the extent that it facilitates the establishment of verification criteria and supports evaluation of its performance \citep{boehm1976quantitative}.
\end{defn}

@@

\noindent \textbf{Reasoning}
Definition is concise and measurable.
Verifiability involves solving the equations right
\citep[p.~23]{Roache1998};
it benefits from rational documentation that systematically shows, with explicit
traceability, how the governing equations are transformed into code.
\oo{Verifiability is sometimes referred to as testability, so I culled some
	testability definitions and added here}  \wss{culled means removed - is that your
	intention?} \oo{culled could also mean  'to choose', so I meant, I choose from some definitions of testability and added to the list}

@@

\begin{defn} \label{Validatability_Selected}
   Validation is the process of evaluating software during or at the end of the development
   process to determine whether it satisfies specified business requirements \citep{softwaretestingfundamentals}.

@@

\begin{defn}
    Software validation checks that the software product satisfies or fits the intended use (high-level checking), i.e., the software meets the user requirements, not as specification artifacts or as needs of those who will operate the software only; but, as the needs of all the stakeholders (such as users, operators, administrators, managers, investors, etc.) \citep{WikipediaVnV}
\end{defn}
\noindent \textbf{Proposed Definition}: Definition~\ref{} \oo{need to discuss this quality, we might just define validatability from validation definitions above? Resources are limited for 'validatability'}.

@@

\noindent \textbf{Proposed Definition}
Definition \ref{reliabilitySelected1} and \ref{reliabilitySelected2} rephrased : 	Probability of failure-free operation of a computer program in a specified environment for a specified time at a specified level of fault tolerance.

@@

\noindent \textbf{Reasoning}
Reliability is defined in a way that the measurement metrics are visible - the specified enviroment, the specified time and the specified level of fault tolerance. Reliability is a critical quality for scientific software, since the results of
computations are meaningless, if they are not dependable.  Reliability is
closely tied to verifiability, since the key quality to verify is reliability,
while the act of verification itself improves reliability.

@@

The Interaction Design
Foundation~\url{https://www.interaction-design.org/literature/topics/usability}
further lists the following desirable outcomes:

@@

\begin{defn} \label{MaintainabilityDefnSelected1}
	The ease with which a software system or component can be modified to correct
	faults, improve performance or other attributes, or adapt to a changed
	environment \citep{IEEEStdGlossarySET1990}. 
\end{defn}
\begin{defn}
	ISO/IEC 25010 refers to maintainability as the degree of effectiveness and
	efficiency with which a product or system can be modified by the intended
	maintainers \citep{ISO/IEC25010}.
\end{defn}
\begin{defn}
	Effort required to locate and fix an error in a program
	\citep{pressman2005software}.
\end{defn}
\begin{defn}
	A set of attributes that bear on the effort needed to make specified
	modifications (which may include corrections, improvements, or adaptations of
	software to environmental changes and changes in the requirements and functional
	specifications) \citep{pfleeger2006software}.
\end{defn}
\begin{defn}
	We will view maintainability as two separate qualities: repairability and
	evolvability. Software is repairable if it allows the fixing of defects; it is
	evolvable if it allows changes that enable it to satisfy new requirements
	\citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn} \label{MaintainabilityDefnSelected2}
	Code possesses the characteristic of maintainability to the extent that it
	facilitates updating to satisfy new requirements or to correct deficiencies
	\citep{boehm2007software}.
\end{defn}

@@

\begin{defn}
	Interoperability is the ability of two or more systems or components to exchange
	information and to use the information that has been exchanged
	\citep{IEEEComputerDictionary1991}.
\end{defn}
\begin{defn}
	\label{InteroperabilitySelected}
	The degree to which two or more systems, products or components can exchange
	information and use the information that has been exchanged
	\citep{ISO/IEC25010}.
\end{defn}
\begin{defn}
	The ability of a system to coexist and cooperate with other systems.
	\citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn}
	The capability to communicate, execute programs, and transfer data among various
	functional units in a manner that requires the user to have little or no
	knowledge of the unique characteristics of those units
	\citep{ISO/IEC/IEEE24765}.
\end{defn}
\begin{defn}Interoperability is a characteristic of a product or system, whose
	interfaces are completely understood, to work with other products or systems,
	present or future, in either implementation or access, without any restrictions
	\citep{AFUL2019}.
\end{defn}
\begin{defn}
	Interoperability is the ability of different information systems, devices and
	applications (‘systems’) to access, exchange, integrate and cooperatively use
	data in a coordinated manner, within and across organizational, regional and
	national boundaries, to provide timely and seamless portability of information
	and optimize the health of individuals and populations globally. Health data
	exchange architectures, application interfaces and standards enable data to be
	accessed and shared appropriately and securely across the complete spectrum of
	care, within all applicable settings and with relevant stakeholders, including
	by the individual \citep{HIMSS2019}.

@@

\begin{defn}
	A quality of the software production process, referring to its efficiency and
	performance \citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn}
	The best definition of the productivity of a process is
	\[\text{Productivity} = \dfrac{\text{Outputs produced by the
			process}}{\text{Inputs consumed by the process}}\] Defining inputs. For
	the software process, providing a meaningful definition of inputs is a
	nontrivial but generally workable problem. Inputs to the software process
	generally comprise labor, computers, supplies, and other support facilities and
	equipment. Defining outputs. The big problem in defining software productivity
	is defining outputs. Here we find a defining delivered source instructions (DSI)
	or lines of code as the output of the software process is totally inadequate,
	and they argue that there are a number of deficiencies in using DSI. However,
	most organizations doing practical productivity measurement still use DSI as
	their primary metric \citep{Boehm1987}. \wss{Is this a direct quote from
		\citet{Boehm1987}?  The sentences seem incomplete?}
\end{defn}
\begin{defn}
	\label{ProductivitySelected}
	Productivity is the amount of output (what is produced) per unit of input
	used. If we can measure the size of the software product and the effort required
	to develop the product, we have:
	\begin{align}
	\text{productivity} = \text{size}/\text{effort}
	\end{align}
	Equation (1) assumes that size is the output of the software production process
	and effort is the input to the process. This can be contrasted with the
	viewpoint of software cost models where we use size as an independent variable
	(i.e., an input) to predict effort which is treated as an output. Equation (1)
	is simple to operationalize if we have a single dominant size measure, for
	example, product size measured in lines of code \citep{Kitchenham2004}.
\end{defn}

@@

\begin{defn}
	\label{ConsistencySelected}
	A specification is consistent to the extent that its provisions do not conflict
	with each other or with governing specifications and objectives. Specifications
	require consistency in several ways \citep{Boehm1984}.
	\begin{itemize}
		\item Internal consistency. Items within the specification do not conflict with
		each other.
		\item External consistency. Items in the specification do not conflict with
		external specifications or entities.
		\item Traceability. Items in the specification have clear antecedents in earlier
		specifications or statements of system objectives.
	\end{itemize}
\end{defn}
\begin{defn}
	Consistency requires that no two or more requirements in a specification
	contradict each other. It is also often regarded as the case where words and
	terms have the same meaning throughout the requirements specifications
	(consistent use of terminology). These two views of consistency imply that
	mutually exclusive statements and clashes in terminology should be avoided
	\citep{ZOWGHI2003}.
\end{defn}
\begin{defn}
	Consistency: 1. the degree of uniformity, standardization, and freedom from
	contradiction among the documents or parts of a system or component 2. software
	attributes that provide uniform design and implementation techniques and
	notations \citep{ISO/IEC/IEEE24765}.
\end{defn}

>>>>

\begin{defn}
  A measure of the ease of installation \citep{smith2015comparing}.\wss{Where
    does this definition come from? Please add a citation.  Is it from
    \citet{McCallEtAl1977}? I thought so originally, but after looking into it,
    I don't think so anymore.}  \oo{@smith you added the definition in here but
    I guess it's from here} \wss{@olu, I was hoping you would verify the source
    of the definition.  Can you please find a pdf of the McCall reference and
    add it to our pub repo.  We should cite the primary source,
    but we can't assume what it says.}
\end{defn}

@@

\wss{The rationale is missing.  Please add.} \oo{Done} The definition captures
the cost or effort - human or machine, and considers not just the initial
installation but uninstallation and re-installation which are all related to
installation activities.  \oo{Done installability}

\subsection{\notdone{Correctness} \oo{owner}}

@@

\begin{defn}
  Testability is the degree to which a requirement is stated in terms that
  permits establishment of test criteria and performance of tests to determine
  whether those criteria have been met \citep{IEEEStdGlossarySET1990}
\end{defn}

\begin{defn}
  Code possesses the characteristic testability to the extent that it
  facilitates the establishment of verification criteria and supports evaluation
  of its performance \citep{boehm1976quantitative}.
\end{defn}

@@

Definition is concise and measurable.  Verifiability involves solving the
equations right \citep[p.~23]{Roache1998}; it benefits from rational
documentation that systematically shows, with explicit traceability, how the
governing equations are transformed into code.  \oo{Verifiability is sometimes
  referred to as testability, so I culled some testability definitions and added
  here} \wss{culled means removed - is that your intention?} \oo{culled could
  also mean 'to choose', so I meant, I choose from some definitions of
  testability and added to the list}

@@

\begin{defn} \label{Validatability_Selected} 
  Validation is the process of evaluating software during or at the end of the
  development process to determine whether it satisfies specified business
  requirements \citep{softwaretestingfundamentals}.
\end{defn}

@@

\begin{defn}
  Software validation checks that the software product satisfies or fits the
  intended use (high-level checking), i.e., the software meets the user
  requirements, not as specification artifacts or as needs of those who will
  operate the software only; but, as the needs of all the stakeholders (such as
  users, operators, administrators, managers, investors, etc.)
  \citep{WikipediaVnV}
\end{defn}

Definition~\ref{} \oo{need to discuss this quality, we might just define
  validatability from validation definitions above? Resources are limited for
  'validatability'}.

@@

Definition \ref{reliabilitySelected1} and \ref{reliabilitySelected2} rephrased:
Probability of failure-free operation of a computer program in a specified
environment for a specified time at a specified level of fault tolerance.

@@

Reliability is defined in a way that the measurement metrics are visible - the
specified enviroment, the specified time and the specified level of fault
tolerance. Reliability is a critical quality for scientific software, since the
results of computations are meaningless, if they are not dependable.
Reliability is closely tied to verifiability, since the key quality to verify is
reliability, while the act of verification itself improves reliability.

@@

The Interaction Design Foundation
\url{https://www.interaction-design.org/literature/topics/usability} further
lists the following desirable outcomes:

@@

\begin{defn} \label{MaintainabilityDefnSelected1} 
  The ease with which a software system or component can be modified to correct
  faults, improve performance or other attributes, or adapt to a changed
  environment \citep{IEEEStdGlossarySET1990}.
\end{defn}

\begin{defn}
  ISO/IEC 25010 refers to maintainability as the degree of effectiveness and
  efficiency with which a product or system can be modified by the intended
  maintainers \citep{ISO/IEC25010}.
\end{defn}

\begin{defn}
  Effort required to locate and fix an error in a program
  \citep{pressman2005software}.
\end{defn}

\begin{defn}
  A set of attributes that bear on the effort needed to make specified
  modifications (which may include corrections, improvements, or adaptations of
  software to environmental changes and changes in the requirements and
  functional specifications) \citep{pfleeger2006software}.
\end{defn}

\begin{defn}
  We will view maintainability as two separate qualities: repairability and
  evolvability. Software is repairable if it allows the fixing of defects; it is
  evolvable if it allows changes that enable it to satisfy new requirements
  \citep{ghezzi1991fundamentals}.
\end{defn}

\begin{defn} \label{MaintainabilityDefnSelected2} 
  Code possesses the characteristic of maintainability to the extent that it
  facilitates updating to satisfy new requirements or to correct deficiencies
  \citep{boehm2007software}.
\end{defn}

@@

\begin{defn}
  Interoperability is the ability of two or more systems or components to
  exchange information and to use the information that has been exchanged
  \citep{IEEEComputerDictionary1991}.
\end{defn}

\begin{defn}
  \label{InteroperabilitySelected}
  The degree to which two or more systems, products or components can exchange
  information and use the information that has been exchanged
  \citep{ISO/IEC25010}.
\end{defn}

\begin{defn}
  The ability of a system to coexist and cooperate with other systems.
  \citep{ghezzi1991fundamentals}.
\end{defn}

\begin{defn}
  The capability to communicate, execute programs, and transfer data among
  various functional units in a manner that requires the user to have little or
  no knowledge of the unique characteristics of those units
  \citep{ISO/IEC/IEEE24765}.
\end{defn}

\begin{defn}
  Interoperability is a characteristic of a product or system, whose interfaces
  are completely understood, to work with other products or systems, present or
  future, in either implementation or access, without any restrictions
  \citep{AFUL2019}.
\end{defn}

\begin{defn}
  Interoperability is the ability of different information systems, devices and
  applications (‘systems’) to access, exchange, integrate and cooperatively use
  data in a coordinated manner, within and across organizational, regional and
  national boundaries, to provide timely and seamless portability of information
  and optimize the health of individuals and populations globally. Health data
  exchange architectures, application interfaces and standards enable data to be
  accessed and shared appropriately and securely across the complete spectrum of
  care, within all applicable settings and with relevant stakeholders, including
  by the individual \citep{HIMSS2019}.

@@

\begin{defn}
  A quality of the software production process, referring to its efficiency and
  performance \citep{ghezzi1991fundamentals}.
\end{defn}

\begin{defn}
  The best definition of the productivity of a process is
  \[\text{Productivity} = \dfrac{\text{Outputs produced by the
        process}}{\text{Inputs consumed by the process}}\] Defining inputs. For
  the software process, providing a meaningful definition of inputs is a
  nontrivial but generally workable problem. Inputs to the software process
  generally comprise labor, computers, supplies, and other support facilities
  and equipment. Defining outputs. The big problem in defining software
  productivity is defining outputs. Here we find a defining delivered source
  instructions (DSI) or lines of code as the output of the software process is
  totally inadequate, and they argue that there are a number of deficiencies in
  using DSI. However, most organizations doing practical productivity
  measurement still use DSI as their primary metric \citep{Boehm1987}. \wss{Is
    this a direct quote from \citet{Boehm1987}?  The sentences seem incomplete?}
\end{defn}

\begin{defn}
  \label{ProductivitySelected}
  Productivity is the amount of output (what is produced) per unit of input
  used. If we can measure the size of the software product and the effort
  required to develop the product, we have:
  \begin{align}
    \text{productivity} = \text{size}/\text{effort}
  \end{align}
  Equation (1) assumes that size is the output of the software production
  process and effort is the input to the process. This can be contrasted with
  the viewpoint of software cost models where we use size as an independent
  variable (i.e., an input) to predict effort which is treated as an
  output. Equation (1) is simple to operationalize if we have a single dominant
  size measure, for example, product size measured in lines of code
  \citep{Kitchenham2004}.
\end{defn}

@@

\begin{defn}
  \label{ConsistencySelected}
  A specification is consistent to the extent that its provisions do not
  conflict with each other or with governing specifications and
  objectives. Specifications require consistency in several ways
  \citep{Boehm1984}.
  \begin{itemize}
  \item Internal consistency. Items within the specification do not conflict
    with each other.
  \item External consistency. Items in the specification do not conflict with
    external specifications or entities.
  \item Traceability. Items in the specification have clear antecedents in
    earlier specifications or statements of system objectives.
  \end{itemize}
\end{defn}

\begin{defn}
  Consistency requires that no two or more requirements in a specification
  contradict each other. It is also often regarded as the case where words and
  terms have the same meaning throughout the requirements specifications
  (consistent use of terminology). These two views of consistency imply that
  mutually exclusive statements and clashes in terminology should be avoided
  \citep{ZOWGHI2003}.
\end{defn}
\begin{defn}
  Consistency: 1. the degree of uniformity, standardization, and freedom from
  contradiction among the documents or parts of a system or component
  2. software attributes that provide uniform design and implementation
  techniques and notations \citep{ISO/IEC/IEEE24765}.
\end{defn}
